
---
title: "Looking for mental health care? West is best"
author: "Jack Kagan and Joccelyn Alvarado"
date: "Sunday December 18th, 2016"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

# Load all necessary packages here
library(ggplot2)
library(dplyr)
library(mosaic)


#
# Albert's Note:
#
# I removed this line: the viewer doensn't work with RMarkdown
# View(Northeast)
```





## Main Article

Write your main article here i.e. the part that will show in print or online.
If you are interested in formatting tips, go to RStudio Menu Bar -> Help -> 
Markdown Quick Reference

Two things that are fair to say about Middlebury is that it is a rigorous academic institution, and that its students hail from all across the country. The whole school represents 47 states and DC, and the middle 50% SAT score for the class of 2020 was between 1940 and 2240. Thus, Middlebury expects a lot out of its students. And this, as is the case with many colleges, contributes to the mental health problem that plagues youth and college students across the country. As for us here in Vermont, during the summer and post-Middlebury, when we go back to our respective homes, the landscape is often very different from the one in rural Vermont, or even in New England. 

We've taken data on over 13,000 mental health services facilities across the U.S. and compiled them into regions - Northeast, South, West, and Midwest. From there, we've compared the amount of facilities in each region that have various treatment options that might be PARTICULARLY relevant to a young person/college student looking for mental health care. 

We got our data from the 2012 National Mental Health Services Survey, which included 15,183 of 16,782 eligible/relevant mental health facilities in the country. 



For all the following graphs, the x axis is the proportion of facilities that offer that particular service, while the y axis is the # of facilities that offer that service.
Green is the south.
Black is the northeast.
Red is the midwest.
Blue is the west. 
```{r, message=FALSE, warning=FALSE, fig.width=8, fig.height=4.5, echo=FALSE}
# Copy this blank code block as many times as you need. Note, if you want to
# code to show in your HTML files, set echo=TRUE above.
# mh2012_puf <- read.csv(~/JJ_FInal_Project/mh2012_puf.csv)


ggplot(data = Northeast, aes(x=sum(MH_intake)/length(MH_intake))) + geom_histogram() + geom_histogram(data = midwest, fill = "red") + geom_histogram(data = South, fill = "green") + geom_histogram(data = West, fill = "blue") + xlab("Mental Health intake facilities by region")


ggplot(data = Northeast, aes(x=sum(Substance_treatmnt)/length(Substance_treatmnt))) + geom_histogram() + geom_histogram(data = midwest, fill = "red") + geom_histogram(data = South, fill = "green") + geom_histogram(data = West, fill = "blue") + xlab("Substance abuse treatment facilities by region")
```
Another issue that's relevant to young people and college students alike is substance abuse. The region with the highest proportion of facilities that offer substance abuse treatment is the west, but this proportion is still under 55%. 

```{r, message=FALSE, warning=FALSE, fig.width=8, fig.height=4.5, echo=FALSE}

ggplot(data = Northeast, aes(x=sum(alldayIP)/length(alldayIP))) + geom_histogram() + geom_histogram(data = midwest, fill = "red") + geom_histogram(data = South, fill = "green") + geom_histogram(data = West, fill = "blue") + xlab("24hr inpatient care proportion by region")

ggplot(data = Northeast, aes(x=sum(PSYCHOSES)/length(PSYCHOSES))) + geom_histogram() + geom_histogram(data = midwest, fill = "red") + geom_histogram(data = South, fill = "green") + geom_histogram(data = West, fill = "blue") + xlab("Schizophrenia, other Psychoses treatment")

ggplot(data = Northeast, aes(x=sum(AUTISM)/length(AUTISM))) + geom_histogram() + geom_histogram(data = midwest, fill = "red") + geom_histogram(data = South, fill = "green") + geom_histogram(data = West, fill = "blue") + xlab("Autism spectrum treatment proportion")
```
We were surprised at the range of autism/autism spectrum treatment across the regions. The midwest leads the pack with just over 45% of facilities having treatment options, but the other end, the south and the west, have only around 35% of facilities with treatment options. 

```{r, message=FALSE, warning=FALSE, fig.width=8, fig.height=4.5, echo=FALSE}
ggplot(data = Northeast, aes(x=sum(ADD_or_CONDUCT)/length(ADD_or_CONDUCT))) + geom_histogram() + geom_histogram(data = midwest, fill = "red") + geom_histogram(data = South, fill = "green") + geom_histogram(data = West, fill = "blue") + xlab("ADD/Conduct disorders treatment proportion")

ggplot(data = Northeast, aes(x=sum(EATING)/length(EATING))) + geom_histogram() + geom_histogram(data = midwest, fill = "red") + geom_histogram(data = South, fill = "green") + geom_histogram(data = West, fill = "blue") + xlab("Eating disorder treatment proportion")
```
Another both low and wide range is on the graph of facilities with eating disorder treatment. The midwest leads with almost 48% percent, while the south trails with just over 36% of facilities providing eating disorder treatment.

```{r, message=FALSE, warning=FALSE, fig.width=8, fig.height=4.5, echo=FALSE}
ggplot(data = Northeast, aes(x=sum(POSTTRAUM1)/length(POSTTRAUM1))) + geom_histogram() + geom_histogram(data = midwest, fill = "red") + geom_histogram(data = South, fill = "green") + geom_histogram(data = West, fill = "blue") + xlab("Post traumatic stress treatment proportion")

ggplot(data = Northeast, aes(sum(treats_LGBTQ)/length(treats_LGBTQ))) + geom_histogram() + geom_histogram(data = midwest, fill = "red") + geom_histogram(data = South, fill = "green") + geom_histogram(data = West, fill = "blue") + xlab("proportion of facilities that treat LGBT clients")
```
Another jarring graph is that of the proportion of facilities that provide service to LGBT clients. The south has a shockingly low proportion here. Even the leading west's proportion is surprisingly low. Yes, as sad as it is, some facilities can refuse to serve because of sexual orientation. 

```{r, message=FALSE, warning=FALSE, fig.width=8, fig.height=4.5, echo=FALSE}
ggplot(Northeast, aes(sum(psychiatric_emergency_walk_in)/length(psychiatric_emergency_walk_in))) + geom_histogram() + geom_histogram(data = midwest, fill = "red") + geom_histogram(data = South, fill = "green") + geom_histogram(data = West, fill = "blue") + xlab("Psychiatric Emergency Walk in availability proportion")

ggplot(Northeast, aes(sum(language_other_than_english)/length(language_other_than_english))) + geom_histogram() + geom_histogram(data = midwest, fill = "red") + geom_histogram(data = South, fill = "green") + geom_histogram(data = West, fill = "blue") + xlab("proportion of facilities w/mental health services in a language other than English")

ggplot(Northeast, aes(sum(PAYASST)/length(PAYASST))) + geom_histogram() + geom_histogram(data = midwest, fill = "red") + geom_histogram(data = South, fill = "green") + geom_histogram(data = West, fill = "blue") + xlab("Proportion of facilities with care for those who cannot pay")
```

We also synthesized all the results for the regions as they relate to each other. For these 12 factors that we designated as releavnt for young adults and college students, a region got 4 points for leading the pack, 3 points for second place, 2 points for 3rd place, and 1 point for 4th place out of the 4. 
By this system, the west got an average score of 3.08.
Midwest: 2.92
Northeast: 2.25
and south came in dead last with a 1.9 average score. 



## Supplementary Materials

Write your extra materials here i.e. extra info for people who want to know
more, but including in the main article would unnecessarily bloat things.

here is the link to the explanation of the survey data we used: https://wwwdasis.samhsa.gov/dasis2/nmhss/2012_nmhss_codebook.pdf
and here is a link to the website
https://wwwdasis.samhsa.gov/dasis2/nmhss.htm

```{r, message=FALSE, warning=FALSE, fig.width=8, fig.height=4.5, echo=FALSE}
# Copy this blank code block as many times as you need. Note, if you want to
# code to show in your HTML files, set echo=TRUE above.

```
